Challenges to the Two-Infall Scenario by Large Stellar Age Catalogs

### 2025-07-07 src/tex/ms.tex
Updates from Overleaf

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/smooth_vs_twoinfall.pdf}
    \caption{Chemical abundance tracks predicted by a two-infall model at $R_{\rm gal}=8\kpc$ (solid curve) versus a model with a smooth SFH at three different radii (dashed curves). The two-infall model adopts the fiducial parameters according to Table \ref{tab:multizone-parameters}, while the smooth SFH adopts the parameters of the ``inside-out'' model of \citet{johnson_stellar_2021}. Both models assume the \yZ{2} yield scale (see Table \ref{tab:yields}). The 2-D histogram shows the number of stars from APOGEE DR17 in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $0\leq|z|<2\kpc$) in each bin of ([Fe/H], [O/Fe]).}
    \label{fig:smooth-vs-twoinfall}
\end{figure}

@@

% \todo{Summary of secular evolution models which claim alpha-bimodality.} 
In contrast to the two-infall scenario, a number of studies have attempted to explain the $\alpha$-bimodality through purely secular processes. 
Using a detailed prescription for radial migration, \citet{schonrich_chemical_2009} produced distinct high- and low-$\alpha$ sequences, but they did not overlap in metallicity space as would be found by later surveys \citep[e.g.,][]{bensby_exploring_2014}. Others have produced a more MW-like $\alpha$-bimdodality using a combination of radial migration and inside-out galaxy growth \citep[e.g.,][]{kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history. \citet{sharma_chemical_2021} and \citet{chen_chemical_2023} showed that a simultaneous decline in the star formation rate and the \aFe ratio is needed to separate the sequences in chemical space under this scenario. \citet{chen_recent_2025} additionally argued that the double sequence in the local age--metallicity relation observed by \citet{nissen_high-precision_2020} can also be explained by a smooth star formation history and inside-out disk growth. However, other GCE models that incorporate both radial migration and smooth, inside-out star formation have failed to predict an $\alpha$-bimodality \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024}. While the $\alpha$-bimodality remains a key piece of evidence for the two-infall scenario, it is not the only proposed explanation.

@@

\begin{table*}
    \centering
    \caption{Median and dispersion in APOGEE parameter uncertainties.}
    \label{tab:uncertainties}
    \begin{tabular}{lcc}
        \hline\hline
        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Median Uncertainty} & \multicolumn{1}{c}{Uncertainty Dispersion ($95\%-5\%$)} \\
        \hline
        $\mathOH$   & $0.019$    & $0.031$ \\
        $\mathFeH$  & $0.0089$   & $0.0060$ \\
        % $\mathOFe$  & $0.019$    & $0.031$ \\
        $\log_{10}(\tau_{\rm NN}/{\rm Gyr})$    & 0.10   & $0.16$ \\
        $\tau_{\rm [C/N]}/{\rm Gyr}$     & 1.4   & 1.8 \\
        \hline
    \end{tabular}
\end{table*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/ofe_feh_best.pdf}
    \caption{Stellar abundance distributions across the disk predicted by our best multi-zone model, with \yZ{2}, ${\rm [X/H]}_{\rm CGM}=-0.7$, $f_\Sigma(R_\odot)=0.25$, $\sigma_{\rm RM8}=3.6\kpc$, and $\eta_\odot=1.8$. Each panel presents a random mass-weighted sample of \num{10000} stellar populations that are drawn from the given ($R_{\rm gal}$, $|z|$) bin and color-coded by age. A Gaussian scatter is applied to each point according to the median age ($\tau_{\rm [C/N]}$) and abundance uncertainties in Table \ref{tab:uncertainties}. The solid and dashed contours enclose 30\% and 80\%, respectively, of the APOGEE data in each region. {\bf Key takeaway:} The predicted distribution from the two-infall model lines up with the APOGEE distribution close to the midplane, but agreement is worse at higher latitudes and in the outer Galaxy.}
    \label{fig:ofe-feh-best}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/lmr_ages.pdf}
    \caption{{\it Top:} The median stellar age as a function of [O/Fe] and [Fe/H] in the Solar annulus ($7\leq R_{\rm gal}<9\kpc$, $0\leq|z|<2\kpc$). The left and center panels plot the output of our best two-infall models, with \yZ{2}, ${\rm [X/H]}_{\rm CGM}=-0.7$, $f_\Sigma(R_\odot)=0.25$, and $\sigma_{\rm RM8}=3.6\kpc$. The model output has been re-sampled to match the APOGEE stellar $|z|$ distribution, and a Gaussian scatter has been applied to the abundances and ages according to Table \ref{tab:uncertainties}. The right panel plots the results from APOGEE using the \citet{leung_variational_2023} age catalog. The contours indicate the density of stars in the [Fe/H]--[O/Fe] plane, and the vertical dashed line denotes the boundary for locally metal-rich (LMR) stars.
    {\it Bottom}: Stellar age distributions in the Solar annulus for all stars (black) and LMR stars (gray). The left and center panels plot the mass-weighted age distributions predicted by the models after forward-modeling age uncertainties, and the right panel plots the \citet{leung_variational_2023} ages for APOGEE stars.
    {\bf Key takeaway:} The two-infall model predicts a fundamentally different age pattern than what is observed, especially for LMR stars.}
    \label{fig:lmr-ages}
\end{figure*}

The two-infall model makes a fundamental prediction about the local stellar age distribution: the most metal-rich stars born in-situ in any region of the Galaxy should come from the metal-rich tail of the first infall sequence, and are therefore older than all of the thin disk stars. As noted in the previous Section, this prediction is apparent in any of the panels in Figure \ref{fig:ofe-feh-best}, especially where $|z|<0.5\kpc$. We investigate this prediction further here.

@@

% \todo{Summary of problem: the two-infall model is boxed in by data.}
Our results show that the APOGEE stellar age catalogs pose several challenges to the two-infall scenario. The [O/Fe] distribution was the observable with the easiest discrepancy to resolve using the parameter space we explored, either by pre-enrichment of the accreted gas or a higher thick-to-thin disk mass ratio. These parameters also allowed us to mitigate, but not completely resolve, the issues with the age--abundance relations caused by dilution. The predicted age distributions could not be reconciled with the data using any of the variations we explored. In the remainder of this section, we discuss other parameters explored by previous studies as well as further extensions to the two-infall model.

@@

Motivated by evidence of a recent period of enhanced star formation \citep[e.g.,][]{ruiz-lara_recurrent_2020}, \citet{spitoni_beyond_2023} and \citet{palla_mapping_2024} extended the two-infall model with a recent $(\lesssim3\Gyr)$ third accretion episode. \citet{spitoni_beyond_2023} argued that the gas dilution resulting from the third infall could explain the population of young, metal-poor stars discovered in {\it Gaia} DR3 \citep{recio-blanco_gaia_2023}, in contrast to the two-infall model of \citet{spitoni_apogee_2021} which predicted a present-day gas metallicity of ${\rm [M/H]}\approx+0.3$ in the Solar neighborhood. \citet{palla_mapping_2024} were similarly motivated by the finding that open clusters with ages $<1\Gyr$ have similar metallicity to those with ages $>3\Gyr$, while the classical two-infall model predicted a steady increase in metallicity over time. However, \citet{palla_mapping_2024} invoke a less massive infall, producing a milder dilution event, than \citet{spitoni_beyond_2023}.

The flat AMR from both stellar age catalogs is even more restrictive of a recent dilution event because of the smaller age uncertainties for younger stars. Figure \ref{fig:yield-outflow} illustrates that the mode and overall distribution of [O/H] has evolved very little over the past $\sim5$ Gyr. Some combination of metal-rich accretion or radial gas flows might reduce the amount of dilution predicted by a recent accretion episode, but the parameter space is tightly constrained by the stellar age data.

@@

\begin{figure}
    \centering
    \includegraphics[width=\onecolumn]{figures/onezone_sfe_hiatus.pdf}
    \caption{Abundance tracks and distributions from one-zone models which experience an efficiency-driven starburst. The blue dashed curve represents the fiducial model that has an exponentially declining infall rate and constant star formation efficiency timescale $\tau_\star=2\,{\rm Gyr}$. The red solid curve plots the output of a model which experiences an enhancement of $\tau_\star$ by a factor of 10, for a duration of 200 Myr, starting at $t=1.4\,{\rm Gyr}$. Both models assume the \yZ{2} yield set, with $y_{\rm Fe}^{\rm Ia}$ reduced by 20\% to better match the model endpoint with the data, and $\eta=1.4$. The greyscale histogram presents the number density of APOGEE stars in the Solar annulus ($7\leq R_{\rm gal}\leq 9\kpc$, $0\leq|z|\leq2\kpc$) in [O/Fe]--[Fe/H] space, and the gray histograms in the marginal panels show the APOGEE stellar abundance distributions.}
    \label{fig:onezone-sfe-hiatus}
\end{figure}

The two-infall model falls into the broader category of GCE models which reproduce the $\alpha$-bimodality by halting or severely limiting star formation for some duration. For the two-infall model, this phase of low star formation immediately precedes the second infall epoch and is due to the relatively short timescale of the first infall epoch. However, as we have shown, the dilution of the ISM resulting from the second infall poses a challenge when comparing to age--abundance data.

A bursty infall history is not the only way to produce a gap in the star formation history. \citet{beane_rising_2024} present a simulated galaxy from the Illustris TNG50 suite that exhibits MW-like bimodality. They argue that the $\alpha$-bimodality is brought on by a brief ($\sim300\,{\rm Myr}$) quiescent period caused by bar formation. The virial mass of their galaxy grows steadily throughout this period, unlike in our two-infall model where the mass grows by a factor of \todo{X} during the 1 Gyr following the second infall.

While our semi-analytic model does not include a Galactic bar, we can explore the effects of a star formation hiatus by artificially boosting the SFE timescale $\tau_\star$ for a period of time. Figure \ref{fig:onezone-sfe-hiatus} illustrates the effect of this SFE-driven hiatus in a one-zone model with an exponentially declining infall rate. During the quiescent period, the [O/Fe] ratio slowly declines due to the delayed contribution of Fe from SNe Ia. Meanwhile, the gas mass continues to increase even as star formation is suppressed. When $\tau_\star$ is lowered at the end of the quiescent period, the high gas mass sparks a moderate star formation burst which causes stellar abundances to ``pile up'' at similar [O/Fe] values. The trough between the high- and low-$\alpha$ sequences results from the star formation returning to pre-quiescence behavior.

Our simple hiatus model offers a few parameters which control the chemical evolution. The onset time of the SFE hiatus controls the position of the high-$\alpha$ sequence: a later onset places the peak at lower [O/Fe]. The duration of the star formation hiatus and the $\tau_\star$ enhancement factor control the strength of the high-$\alpha$ peak.

The parameters of the SFE hiatus in Figure \ref{fig:onezone-sfe-hiatus} were chosen to match the APOGEE stellar [O/Fe] distribution as closely as possible. However, there are some differences in detail, such as the dearth of stars at $\mathOFe\approx+0.35$ due to the star formation hiatus. We intend this model to illustrate another path to reproducing the $\alpha$-bimodality. Most of the high-$\alpha$ stars present in the Solar annulus have likely migrated from the inner Galaxy, where perhaps this SFE-driven hiatus was concentrated.

@@

We have compared the predictions of the two-infall scenario against abundance data from APOGEE DR17 supplemented with age estimates using two different methods. We ran multi-zone GCE models at two different yield scales with prescriptions for radially-dependent outflows and stellar migration. While the two-infall scenario can explain the distinct high- and low-$\alpha$ sequences, it faces challenges in matching the age--abundance structure of the full disk. We explored multiple parameter modifications to bring the model predictions closer to the data, including the yield scale, radial migration strength, metallicity of the accreted gas, thick-to-thin disk mass ratio, and the SN Ia DTD. Our conclusions are as follows:

\begin{itemize}
    \item The large quantity of pristine gas accreted in the Solar neighborhood during the second infall phase rapidly dilutes the ISM metallicity by $\sim0.5$ dex. Models with low nucleosynthetic yields (\yZ{1}) remain at sub-Solar metallicity until the present day, in stark contrast to the observed local age--metallicity relation. Models with higher yields and outflows approach the present-day metallicity more rapidly, while pre-enriched infall can reduce the magnitude of the dilution (but not eliminate it entirely). %This issue can be mitigated with higher yields and outflows, but not by pre-enriched infall or radial migration.
    \item For metal-rich stars, the two-infall model predicts a sharp divide in the stellar age distribution between the thick and thin disk populations. In contrast, the data show a smooth gradient between the oldest and youngest stars, with most of the metal-rich stars having intermediate ages.
    \item The ``turn-over'' in the evolution of [O/Fe] following the second infall produces a double-peaked low-$\alpha$ sequence with a fundamentally different abundance structure than observed, especially for models with higher yields. A low yield set (\yZ{1}) coupled with lower outflows, or pre-enrichment of the infalling gas, can bring the stellar [O/Fe] distributions more in line with the data. The parameter space is nonetheless restricted by the need to suppress this feature.
    \item Our models predict that the MDF evolves to higher metallicity over time throughout the disk. This contrasts with the APOGEE data, which show very little change in the mode over the past $\sim6-8$ Gyr.
    % \item The equilibrium scenario of chemical evolution, if correct, places stricter limits on the two-infall model than other evolutionary models.
    \item Adjusting the strength of radial migration naturally affects the scatter in the age--metallicity relation but does not affect the median age.
\end{itemize}

@@

\todo{Personal acknowledgements.}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/smooth_vs_twoinfall.pdf}
    \caption{Chemical abundance tracks predicted by a two-infall model at $R_{\rm gal}=8\kpc$ (solid curve) versus a model with a smooth SFH at three different radii (dashed curves). The two-infall model adopts the fiducial parameters according to Table \ref{tab:multizone-parameters}, while the smooth SFH adopts the parameters of the ``inside-out'' model of \citet{johnson_stellar_2021}. Both models assume the \yZ{2} yield scale (see Table \ref{tab:yields}). The 2-D histogram shows the number of stars from APOGEE DR17 in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $0\leq|z|<2\kpc$) in each bin of ([Fe/H], [O/Fe]).}
    \label{fig:smooth-vs-twoinfall}
\end{figure}

@@

% \todo{Summary of secular evolution models which claim alpha-bimodality.} 
In contrast to the two-infall scenario, a number of studies have attempted to explain the $\alpha$-bimodality through purely secular processes. 
\citet{schonrich_chemical_2009} first showed that radial migration can produce an $\alpha$-bimodality through a superposition of chemical tracks, although they did not directly compare against MW data. Others have produced a more MW-like $\alpha$-bimdodality using a combination of radial migration and inside-out galaxy growth \citep[e.g.,][]{kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history. \citet{sharma_chemical_2021} and \citet{chen_chemical_2023} showed that a simultaneous decline in the star formation rate and the \aFe ratio is needed to separate the sequences in chemical space under this scenario. \citet{chen_recent_2025} additionally argued that the double sequence in the local age--metallicity relation observed by \citet{nissen_high-precision_2020} can also be explained by a smooth star formation history and inside-out disk growth. However, other GCE models that incorporate both radial migration and smooth, inside-out star formation have failed to predict an $\alpha$-bimodality \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024}. While the $\alpha$-bimodality remains a key piece of evidence for the two-infall scenario, it is not the only proposed explanation.

@@

\begin{table}
    \centering
    \caption{Median and dispersion in APOGEE parameter uncertainties.}
    \label{tab:uncertainties}
    \begin{tabular}{l|cc}
        \hline\hline
        & Median & Uncertainty \\
        Parameter & Uncertainty & Dispersion ($95\%-5\%$) \\
        \hline
        $\mathOH$   & $0.019$    & $0.031$ \\
        $\mathFeH$  & $0.0089$   & $0.0060$ \\
        % $\mathOFe$  & $0.019$    & $0.031$ \\
        $\log_{10}(\tau_{\rm NN}/{\rm Gyr})$    & 0.10   & $0.16$ \\
        $\tau_{\rm [C/N]}/{\rm Gyr}$     & 1.4   & 1.8 \\
        \hline
    \end{tabular}
\end{table}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/ofe_feh_best.pdf}
    \caption{Stellar abundance distributions across the disk predicted by our best multi-zone model, with \yZ{2}, ${\rm [X/H]}_{\rm CGM}=-0.7$, $f_\Sigma(R_\odot)=0.25$, $\sigma_{\rm RM8}=3.6\kpc$, and $\eta_\odot=1.8$. Each panel presents a random mass-weighted sample of \num{10000} stellar populations that are drawn from the given ($R_{\rm gal}$, $|z|$) bin and color-coded by age. A Gaussian scatter is applied to each point according to the median age ($\tau_{\rm [C/N]}$) and abundance uncertainties in Table \ref{tab:uncertainties}. The solid and dashed contours enclose 30\% and 80\%, respectively, of the APOGEE data in each region. {\bf Key takeaway:} The predicted distribution from the two-infall model lines up with the APOGEE distribution close to the midplane, but agreement is worse at higher latitudes and in the outer Galaxy.}
    \label{fig:ofe-feh-best}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/lmr_ages.pdf}
    \caption{{\it Top:} The median stellar age as a function of [O/Fe] and [Fe/H] in the Solar annulus ($7\leq R_{\rm gal}<9\kpc$, $0\leq|z|<2\kpc$). The left and center panels plot the output of our best two-infall models, with \yZ{2}, ${\rm [X/H]}_{\rm CGM}=-0.7$, $f_\Sigma(R_\odot)=0.25$, and $\sigma_{\rm RM8}=3.6\kpc$. The model output has been re-sampled to match the APOGEE stellar $|z|$ distribution, and a Gaussian scatter has been applied to the abundances and ages according to Table \ref{tab:uncertainties}. The right panel plots the results from APOGEE using the \citet{leung_variational_2023} age catalog. The contours indicate the density of stars in the [Fe/H]--[O/Fe] plane, and the vertical dashed line denotes the boundary for locally metal-rich (LMR) stars.
    {\it Bottom}: Stellar age distributions in the Solar annulus for all stars (black) and LMR stars (gray). The left and center panels plot the mass-weighted age distributions predicted by the models after forward-modeling age uncertainties, and the right panel plots the \citet{leung_variational_2023} ages for APOGEE stars.
    {\bf Key takeaway:} The two-infall models predict a fundamentally different age pattern than what is observed, especially for LMR stars.}
    \label{fig:lmr-ages}
\end{figure*}

The two-infall scenario makes a fundamental prediction about the local stellar age distribution: the most metal-rich stars born in-situ in any region of the Galaxy should come from the metal-rich tail of the first infall sequence, and are therefore older than all of the thin disk stars. As noted in the previous Section, this prediction is apparent in any of the panels in Figure \ref{fig:ofe-feh-best}, especially where $|z|<0.5\kpc$. We investigate this prediction further here.

@@

% \todo{Summary of problem: the two-infall model is boxed in by data.}
% Our results show that the APOGEE stellar age catalogs pose several challenges to the two-infall scenario. The [O/Fe] distribution was the observable with the easiest discrepancy to resolve using the parameter space we explored, either by pre-enrichment of the accreted gas or a higher thick-to-thin disk mass ratio. These parameters also allowed us to mitigate, but not completely resolve, the issues with the age--abundance relations caused by dilution. The predicted age distributions could not be reconciled with the data using any of the variations we explored.

In this work, we have highlighted a few predictions of the two-infall scenario and compared against trends from large stellar age catalogs. We have explored a wide range of parameter space, but our investigation is not intended to be exhaustive. In the remainder of this Section, we discuss other parameters explored by previous studies as well as further extensions to the two-infall model.

@@

Motivated by evidence of a recent period of enhanced star formation across the MW disk \citep{ruiz-lara_recurrent_2020}, \citet{spitoni_beyond_2023} extended the two-infall scenario with a recent $(\lesssim3\Gyr)$ third accretion episode. In contrast to the two-infall model of \citet{spitoni_apogee_2021}, which predicted a present-day gas metallicity of ${\rm [M/H]}\approx+0.3$ in the Solar neighborhood, \citet{spitoni_beyond_2023} argued that the gas dilution resulting from the third infall could explain a population of young stars with slightly sub-Solar abundances discovered in {\it Gaia} DR3 \citep{recio-blanco_gaia_2023}.
% \citet{palla_mapping_2024} were similarly motivated by the finding that open clusters with ages $<1\Gyr$ have similar metallicity to those with ages $>3\Gyr$, while the classical two-infall model predicted a steady increase in metallicity over time. However, \citet{palla_mapping_2024} invoke a less massive infall, producing a milder dilution event, than \citet{spitoni_beyond_2023}.

Similarly, \citet{palla_mapping_2024} invoked a late-time accretion episode to explain largely age-independent metallicites in open clusters across the disk. Open clusters trace younger populations: most of their clusters are $\lesssim2\Gyr$ old. A third infall can help explain trends within this limited age range, but when compared to the full range of stellar age data, it faces the same challenges as the second infall. Successive epochs of dilution and re-enrichment will always struggle to explain chemical abundances that remain nearly constant with age over a $\sim10\Gyr$ period. As discussed in Section \ref{sec:age-abundance}, pre-enriched infall could reduce the magnitude of the dilution from a recent accretion episode, but the stellar age data place tight constraints on the parameter space.

@@

\begin{figure}
    \centering
    \includegraphics[width=\onecolumn]{figures/onezone_sfe_hiatus.pdf}
    \caption{Abundance tracks and distributions from one-zone models which experience an efficiency-driven starburst. The blue dashed curve represents the fiducial model that has an exponentially declining infall rate with timescale $\tau_{\rm in}=15\Gyr$, and constant star formation efficiency timescale $\tau_\star=2\,{\rm Gyr}$. The red solid curve plots the output of a model which experiences an enhancement of $\tau_\star$ by a factor of 10, for a duration of 200 Myr, starting at $t=1.4\,{\rm Gyr}$. Both models assume the \yZ{2} yield set, with $y_{\rm Fe}^{\rm Ia}$ reduced by 20\% to better match the model endpoint with the data, and $\eta=1.4$. The grayscale histogram presents the number density of APOGEE stars in the Solar annulus ($7\leq R_{\rm gal}\leq 9\kpc$, $0\leq|z|\leq2\kpc$) in [O/Fe]--[Fe/H] space, and the gray histograms in the marginal panels show the APOGEE stellar abundance distributions.}
    \label{fig:onezone-sfe-hiatus}
\end{figure}

The two-infall scenario falls into the broader category of GCE models that reproduce the $\alpha$-bimodality by halting or severely limiting star formation for some duration. For the two-infall model, this phase of low star formation lasts for a few Gyr in between the $e$-folding timescale of the first infall and the onset of the second. However, as we have shown in Section \ref{sec:age-abundance}, the resulting dilution of the ISM is in tension with the observed stellar age--metallicity relation.

An accretion-driven star formation history is not the only way to induce a hiatus in star formation. Investigating a simulated galaxy from the Illustris TNG50 suite \citep{pillepich_first_2019,nelson_first_2019,nelson_illustristng_2019}, \citet{beane_rising_2024} argued that its MW-like $\alpha$-bimodality is the result of a brief ($\sim300\,{\rm Myr}$) quiescent period caused by the formation of a bar. In this scenario, the newly-formed bar funnels gas into the galaxy's central black hole, prompting a burst of AGN activity which suppresses star formation. Unlike our two-infall model, where the gas mass suddenly grows by a factor of $\sim5$ during the 1 Gyr following the second infall, the mass of their galaxy grows slowly during the quiescent period, indicating a dearth of major accretion events.

While our semi-analytic model does not include a Galactic bar, we can induce a star formation hiatus by artificially boosting the SFE timescale $\tau_\star$ for a period of time. Figure \ref{fig:onezone-sfe-hiatus} illustrates the effect of this efficiency-driven hiatus in a one-zone GCE model with an exponentially declining infall rate ($\tau_{\rm in}=15\Gyr$). During the quiescent period, the [O/Fe] ratio slowly declines due to the delayed contribution of Fe from SNe Ia. Meanwhile, the gas mass continues to increase even as star formation is suppressed. When $\tau_\star$ is lowered at the end of the quiescent period, it sparks a moderate star formation burst which causes stellar abundances to ``pile up'' at similar [O/Fe] values. The trough between the high- and low-$\alpha$ sequences results from the star formation returning to pre-quiescence behavior. The onset time of the SFE hiatus controls the position of the high-$\alpha$ sequence\textemdash a later onset places the peak at lower [O/Fe]\textemdash while the duration and enhancement factor control the strength of the high-$\alpha$ peak.

The parameters of the SFE hiatus in Figure \ref{fig:onezone-sfe-hiatus} were chosen to match the APOGEE stellar [O/Fe] distribution as closely as possible. As such, we start the hiatus at $t=1.4\Gyr$ to match the high-$\alpha$ locus, even though this is $\sim3\Gyr$ before the median age of the thick disk as measured by \citet{pinsonneault_apokasc-3_2025}. This hiatus model illustrates an alternative scenario for reproducing the $\alpha$-bimodality, but a full multi-zone treatment is beyond the scope of this work.
% We intend this model to illustrate another path to reproducing the $\alpha$-bimodality. Most of the high-$\alpha$ stars present in the Solar annulus have likely migrated from the inner Galaxy, where perhaps this SFE-driven hiatus was concentrated.

@@

We have compared the predictions of the two-infall scenario against abundance data from APOGEE DR17 supplemented with age estimates using two different methods. We ran multi-zone GCE models at two different yield scales with prescriptions for radially-dependent outflows and stellar migration. While the two-infall scenario can explain the distinct high- and low-$\alpha$ sequences, it faces challenges in matching the age--abundance structure of the full disk. We explored multiple parameter modifications to bring the model predictions closer to the data, including the yield scale, radial migration strength, metallicity of the accreted gas, thick-to-thin disk mass ratio, and the SN Ia DTD. We found that our two-infall models faced the following challenges when compared against the stellar age catalogs:

\begin{itemize}
    \item By the model's design, the Galaxy undergoes a period of major, rapid dilution and subsequent re-enrichment. Models at a lower yield scale and weaker outflows predict slower re-enrichment. By contrast, the stellar age--metallicity relation is roughly constant over the past $\sim10\Gyr$. Pre-enriching the accreted gas can reduce the magnitude of the dilution, but not eliminate it entirely.
    \item Due to this re-enrichment, our models predict that the MDF evolves to higher metallicity over time throughout the disk. This contrasts with the APOGEE data, which show very little change in the mode of the MDF over the past $\sim6-8$ Gyr.
    \item For metal-rich stars, the two-infall model predicts a sharp divide in the stellar age distribution between the thick and thin disk populations. In contrast, the data show a smooth gradient between the oldest and youngest stars, with most of the metal-rich stars having intermediate ages. This discrepancy could not be reconciled using any of the model variations we explored.
    \item The ``turn-over'' in the evolution of [O/Fe] following the second infall produces a double-peaked low-$\alpha$ sequence with a fundamentally different abundance structure than observed, especially for models with higher yields. A low yield set (\yZ{1}) coupled with lower outflows, or pre-enrichment of the infalling gas, can bring the stellar [O/Fe] distributions more in line with the data. The parameter space is nonetheless restricted by the need to suppress this feature.
    % \item Adjusting the strength of radial migration naturally affects the scatter in the age--metallicity relation but does not affect the median age.
\end{itemize}

We note that we have not found a GCE model which is capable of reproducing all of the above observations. The ``inside-out'' models of \citet{johnson_stellar_2021} are better able to predict a slowly evolving metallicity gradient, but they do not predict an $\alpha$-bimodality. Even a smoothly-evolving SFH struggles to reproduce the flat age--metallicity relation at the lower empirical yield scale \citep{johnson_milky_2024}. We emphasize, though, that by its very construction the two-infall scenario trades success at reproducing certain observations for failure at others.

@@

\todo{Personal acknowledgements.} We are grateful to Prof.\ David Weinberg for his thoughtful comments and input on this paper.
